### 2023-06-14 definitions.tex
up

%The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).

% \textcolor{red}{Before defining WSTS, need a definition of TS and WQO}

@@

 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
%
%
\iffalse
%
\begin{lemma}
(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
\end{lemma}

\begin{proof}
Consider an infinite sequence and the set $M = \{i \in N \mid \forall j > i ~ x_i \not\leq x_j \}$. $M$ cannot
be infinite, otherwise it would lead to an infinite subsequence contradicting the wqo
hypothesis. Thus, $M$ is bounded and any $x_i$ with $i$ beyond $M$ can start an infinite
increasing subsequence.
\end{proof}
%
\fi
%
Wqo admits many other equivalent formulations. 
% For instance:  $(X,\leq)$ is a wqo f and only if any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$) if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of upward-closed subsets in $X$ eventually stabilizes, if and only if any upward-closed subset $U \subseteq X$ has a finite basis if and only if $(X,\leq)$ is well-founded (i.e. it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
%
\iffalse
%

\begin{lemma}{(Higman [40])} 
If $\leq$ is a wqo then any upward-closed $I$ has a
finite basis.
\end{lemma}

% \textcolor{red}{Expliquer ce que c'est une base d'abords.}

\begin{proof}
The set of minimal elements of $I$ is a basis because $\leq$ is well-founded. It
only contains a finite number of non-equivalent elements otherwise they would make
an infinite sequence contradicting the wqo assumption.
\end{proof}

\begin{lemma}  \label{upward-closed stablizes}
If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
upward-closed sets eventually stabilizes, i.e. there is a $k \in N$ such that 
$I_k = I_{k+1} = I_{k+2} = \cdots $.
\end{lemma}

\begin{proof}
Assue we have a counter-example.
We extract an infinite subsequence where
inclusion is strict: $I_{n_0} \subsetneq I_{n_1} \subsetneq I_{n_2} \cdots$. Now, for any $i>0$, we can pick some $x_i \in I_{n_i} \setminus I_{n_{i-1}}$. The well-quasi-ordering hypothesis means that the infinite sequence of $x_i$'s
contains an increasing pair $x_i \leq x_j$ for some $i<j$. Because $x_i$ belongs to an upward-
closed set $I_{n_i}$ we have $x_j \in I_{n_i} $, contradicting $x_j \not\in I_{n_{ j - 1}}$.
\end{proof}
%
\fi
%
Quasi-orderings that have no 
%infinite antichains
infinite subset of mutually incomparable elements (antichains)
 enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

@@

\alain{supprimer ce qui est inutile}

@@

\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

@@

% \textcolor{red}{Define 'has effective pred-basis'. Maybe it should be included in WSTS definition, maybe it can be another def. I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}
\iffalse
\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
If $\mathscr{S}$ is an WSTS and $U \subseteq S$ is an upward-closed set of states, then $\pred^*(U )$ is upward-
closed.
\end{proposition}
%
Proof. Assume $s \in \pred^* (U )$. Then $s \rightarrow^* t$ for some $t \in U $. If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$. Then $t' \in U$ and $s' \in \pred^*(U )$.
\fi

@@

%\alain{étrange ces deux propositions, l'une suffirait}
%%%%%%%%%%%%%%%%%
\iffalse
%%%%%%%%%%%%%
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
%%%%%%%%%%%
\fi
%%%%%%%%%%

Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$. \\

With the effective pred-basis hypothesis, we obtain:

>>>>

@@

\mathieu{Mentionner le qo utilisé dans l'example~\ref{Example}}

 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
%
%
%
Wqo admits many other equivalent formulations. 
% For instance:  $(X,\leq)$ is a wqo f and only if any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$) if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of upward-closed subsets in $X$ eventually stabilizes, if and only if any upward-closed subset $U \subseteq X$ has a finite basis if and only if $(X,\leq)$ is well-founded (i.e. it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
%
%
Quasi-orderings that have no 
%infinite antichains
infinite subset of mutually incomparable elements (antichains)
 enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

@@

@@

\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est un WSTS}

@@

@@

\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est effectif}

Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$. 
With the effective pred-basis hypothesis, we obtain:
